### 2017-05-10 presentation.tex
Fill in ZINB slides tweak intro paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (10/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND THEMES
%----------------------------------------------------------------------------------------

\documentclass{beamer}

\mode<presentation> {

	% The Beamer class comes with a number of default slide themes
	% which change the colors and layouts of slides. Below this is a list
	% of all the themes, uncomment each in turn to see what they look like.

	\usetheme{default}
	%\usetheme{AnnArbor}
	%\usetheme{Antibes}
	%\usetheme{Bergen}
	%\usetheme{Berkeley}
	%\usetheme{Berlin}
	%\usetheme{Boadilla}
	%\usetheme{CambridgeUS}
	%\usetheme{Copenhagen}
	%\usetheme{Darmstadt}
	%\usetheme{Dresden}
	%\usetheme{Frankfurt}
	%\usetheme{Goettingen}
	%\usetheme{Hannover}
	%\usetheme{Ilmenau}
	%\usetheme{JuanLesPins}
	%\usetheme{Luebeck}
	%\usetheme{Madrid}
	%\usetheme{Malmoe}
	%\usetheme{Marburg}
	%\usetheme{Montpellier}
	%\usetheme{PaloAlto}
	%\usetheme{Pittsburgh}
	%\usetheme{Rochester}
	%\usetheme{Singapore}
	%\usetheme{Szeged}
	%\usetheme{Warsaw}

	% As well as themes, the Beamer class has a number of color themes
	% for any slide theme. Uncomment each of these in turn to see how it
	% changes the colors of your current slide theme.

	%\usecolortheme{albatross}
	\usecolortheme{beaver}
	%\usecolortheme{beetle}
	%\usecolortheme{crane}
	%\usecolortheme{dolphin}
	%\usecolortheme{dove}
	%\usecolortheme{fly}
	%\usecolortheme{lily}
	%\usecolortheme{orchid}
	%\usecolortheme{rose}
	%\usecolortheme{seagull}
	%\usecolortheme{seahorse}
	%\usecolortheme{whale}
	%\usecolortheme{wolverine}

	%\setbeamertemplate{footline} % To remove the footer line in all slides uncomment this line
	%\setbeamertemplate{footline}[page number] % To replace the footer line in all slides with a simple slide count uncomment this line

	%\setbeamertemplate{navigation symbols}{} % To remove the navigation symbols from the bottom of all slides uncomment this line
}

\usepackage{sidecap}%[capbesideposition=inside, facing=yes,capbesidesep=quad]
\usepackage{graphicx} % Allows including images
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables
%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[Four Models]{Four Models for Severe Weather Deaths} % The short title appears at the bottom of every slide, the full title is only on the title page

\author{Jason Michaels (jam521), Niko Paulson (ndp32), \\Miranda Seitz-McLeese (mgs85) } % Your name
%\institute[UCLA] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
%{
%	University of California \\ % Your institution for the title page
%	\medskip
%	\textit{john@smith.com} % Your email address
%}
\date{May 11, 2017} % Date, can be changed to a custom date

\begin{document}

	\begin{frame}
		\titlepage % Print the title page as the first slide
	\end{frame}

	\begin{frame}
		\frametitle{Overview} % Table of contents slide, comment this block out to remove it
		\tableofcontents % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
	\end{frame}

	%----------------------------------------------------------------------------------------
	%	PRESENTATION SLIDES
	%----------------------------------------------------------------------------------------
	\section{Introduction} 
	\begin{frame}
		\frametitle{Introduction: Goals and Motivation}
		Understanding how and at what rate severe weather events become lethal in the United States has tremendous health impacts. Understanding the distribution of death counts can help meteorologists better predict outcomes.
	\end{frame}
	\begin{frame}
		\frametitle{Introduction: Data}
		The data for this project was taken from the NOAA severe weather events dataset. 
		In its original form this dataset includes a variety of measurements for all forms of severe weather in the United States dating from 1950. In the interest of time we restricted the dataset to:
		\begin{itemize}
			\item Deaths directly attributable to Tornadoes and Flash floods 
			\item Occurred between 1996-2016
		\end{itemize}
	\end{frame}
	\begin{frame}
		\frametitle{Introduction: Zero Inflated Models}
			Fortunately, the vast majority of severe weather events in the United States involve no deaths, therefore we wanted to also account for the possibility of structural zeros, therefore we also considered zero inflated variants. A zero inflated model is a combination of the original distribution and a point mass at zero. If the original distribution is $p(x)$ then $$z(x) = \begin{cases}
			\sigma + (1-\sigma)*p(0) &x=0\\
			(1-\sigma)*p(x) &x\ne0
			\end{cases}$$
	\end{frame}
	\begin{frame}
		\frametitle{Introduction: Four Models}
				In this paper we compare four possible models for the deaths: 
				\begin{enumerate}
					\item Poisson
					\item Negative Binomial
					\item Zero Inflated Poisson
					\item Zero Inflated Negative Binomial
				\end{enumerate}
	\end{frame}

	\section{Negative Binomial Model}
	\begin{frame}
		\frametitle{Negative Binomial}
		% Slide one
	\end{frame}
	\begin{frame}
		\frametitle{Negative Binomial}
		% Slide two
	\end{frame}

	\section{Poisson Model}
	\begin{frame}
		\frametitle{Poisson Model}
		% Slide one
	\end{frame}
	\begin{frame}
		\frametitle{Poisson Model}
		% Slide two
	\end{frame}

	\section{Zero Inflated Poisson Model}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}

	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}

	\end{frame}

	\section{Zero Inflated Negative Binomial Model}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model}
		The Zero Inflated Negative Binomial (ZINB) model has three parameters $\sigma,$ the probability of a structural zero, and $p,r$ the usual negative binomial parameters. 
		For a single $X$ the probability density is: 
		$$p(X|\sigma, p, r) = \sigma I_{X=0}(X) + (1-\sigma)\frac{\Gamma(r+X)}{\Gamma(r)X!}p^X(1-p)^r$$
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model}
		We took the uniform priors on $\sigma$ and $p$ and the non-informative prior $r^{-1/2}$ on $r$.
		This gave a posterior of:
		\scriptsize$$p(r,\sigma, p|X)\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}\prod_{i=1}^N\left(\frac{\Gamma(r+X_i)}{r^{1/2}\Gamma(r)}\right)$$
		\normalsize This factors into nothing. Therefore we used the Metropolis Hastings Algorithm with proposal distributions of truncated normals.
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model:Results}
		\begin{itemize}
			\item Curse of Dimensionality; very slow convergence.
			\item High autocorrelation
			\item Low confidence in results
		\end{itemize}
		% parameter table goes here
	\end{frame}

	\section{Conclusions}
	\begin{frame}
		\frametitle{Conclusions: Model Evaluation Table}
		% Waic dic table goes here
	\end{frame}
	\begin{frame}
		\frametitle{Conclusions: Interpreting the Best Model}
		% Waic dic table goes here
	\end{frame}
	\begin{frame}
		\frametitle{Conclusions: Future Work}
		% List of things we might have done if we'd had time
		\begin{enumerate}
			\item{Re-parameterize the ZINB model/look into other techniques to improve convergence.}
			\item{See if the results hold on other extreme weather events.}
		\end{enumerate}
	\end{frame}
	\begin{frame}
		\frametitle{Conclusions}
		Questions?
	\end{frame}

\end{document}